---
title: "Green Building(P#1)"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

I build a predictive model of the rent by using greenbuilding data.
The data contains information of 7,894 rental properties including rent per square foot, green rating related information, and other features.

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(mosaic)
library(tidyverse)
library(gamlr) 
```
```{r}
gbldg = read.csv("../data/greenbuildings.csv")
nrow(gbldg)
```

### The data cleansing

There are some missing data in the data set. Missing data can make an error to build a model, so delete the data rows which have missing data.

```{r}
length(which(complete.cases(gbldg)=="FALSE")) 
gbldg[which(complete.cases(gbldg)=="FALSE"),"empl_gr"]
gbldg = na.omit(gbldg)
nrow(gbldg)
```

As we can see, 74 rows have omitted values in 'empl_gr' variable. After delete the data, there are 7820 rental properties' information.

### Other features of data

The green rating variable is an indicator for LEED or EnergyStar certificate. However, there are only `r length(which(gbldg$LEED==1)) ` LEED certificates, while `r length(which(gbldg$Energystar==1)) ` Energy Star certificates. So, I will collapse LEED and EnergyStar into a single "green certified".

The data in the same cluster have the same features ; 
> which are empl_gr, cd_total_07, hd_total07, total_dd_07, Precipitation, Gas_Costs, Electricity_Costs, cluster_rent 

Because these variables are the features of the clustered area.

## Build a predictive model

I build a model by using stepwise selection. 
First, I make an initial model for proceeding a stepwise selection.
Lasso regreesion can make me save the time to build a model, I apply a lasso regression. Then I will step forward to the model with interaction variables.

+ There was an error to build second order sparse matrices for the Lasso regression with interaction variables. Thus, I build a Lasso model without interaction, then build a predictive model by stepwise selection from the Lasso model.

### LASSO regreesion for initial model

First, I make a sparse matrix without CS_PropertyID, cluster, LEED and Energystar variables, because CS_PropertyID and cluster variables are a kind of name, the numerical size of the data have no meaning, and I use green rating vairable instead of LEED and Energystar.

```{r}
gb_x = sparse.model.matrix(Rent ~ .-CS_PropertyID - cluster - 
      LEED - Energystar, data = gbldg)[,-1]
gb_y = gbldg$Rent
```

The lasso regression result shows that the model is consists of 15 variables including size, green_rating, renovated and so on.

```{r}
gblasso = gamlr(gb_x, gb_y)
coef(gblasso)
```

### Stepwise selection

My initial model is

```{r}
lm0 = lm(Rent ~ size + empl_gr + leasing_rate + age +
           renovated + class_a + class_b +
           green_rating + net + amenities + 
           cd_total_07 + hd_total07 +
           Gas_Costs + Electricity_Costs + 
           cluster_rent, data = gbldg)
```

Then, I use step function to step forward from my initial model to the model with interaction variables.

```{r message=FALSE, warning=FALSE, paged.print=FALSE, results = "hide"}
lm_st = step(lm0, scope = ~ (size + empl_gr + leasing_rate + stories + age + 
                             renovated + class_a + class_b + green_rating + net + amenities +
                             cd_total_07 + hd_total07 + total_dd_07 + Precipitation + 
                             Gas_Costs + Electricity_Costs + cluster_rent)^2)

```

The result shows that the final model is 
```{r}
summary(lm_st)
```

The AIC of the initial model is `r min(AICc(gblasso))` , and the final model's AIC is 34,510.

## The change in rental income associated with green certification

The coefficient of green_rating is 2.429, and this variable is statistically significant at 1% level. 
The coefficient shows that green certification is associated with 2.429 additional dollars per square foot per year, holding all else fixed.

## The relationship of the effect of green certification and others

The model shows that there are interaction variables between green_rating variable and amenities variable. This means that the green certification with amenities near the building is associated with 0.1( = 2.429 -2.3) additional dollars per square foot per year, holding all else fixed.


