---
title: "Grocery purchases(P#3)"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
library(igraph) 
library(arules) 
library(arulesViz)
```

This is an association rule mining with groceries data.  
The data is a list of shopping basket, which is divided by "," and "enter".  
So, I split items in one's shopping basket for an association rule mining.

```{r}
groc_raw = readLines("../data/groceries.txt")
head(groc_raw, 3)
# transform for an association rule mining
groc = lapply(groc_raw,strsplit,split=",") # split by ','
groc = lapply(groc, unique) # delete repeated values
groc = lapply(groc, unlist) # list to vector
head(groc, 3)
```

## Association rule mining

Now, I run an association rule analysis by using 'arules' package.  
First, I make a transaction table, and see what are the most frequent items.

```{r}
# transaction table
groctrans = as(groc, "transactions") 
itemFrequencyPlot(groctrans, topN=20)
```

Next, make association rules, with minimum support 0.01, and minimum confidence 0.1, at least one item, and at most 10 items. There are **435** rules generated. 

```{r}
cart = apriori(groctrans, parameter=list(support=.01, confidence=.1, minlen=1, maxlen=10))
```

Focus on some meaningful association rules. I extract association rules with confidence being 0.02 or higher. There are 232 rules with the condition.  
Plot 50 association rules with highest lift valuses. 
One of strong red associations is from 'citrus fruit' to 'root vegetables'.  

```{r}
sub1 = subset(cart, subset = confidence > 0.2)
plot(head(sub1, 50, by='lift'), method='graph')
```

Here are 5 rules with the highest lift value.  
People who buy 'citrus fruit' and 'other vegetables' tend to buy 'root vegetables'. They might be people who love fruits and vegetables.   
People who buy 'beef' tend to buy 'root vegetables'. They might want to make garnish with root vegetables for their steak.

```{r}
inspect(sort(sub1, by = "lift")[1:5]) 
```

People mostly buy 'whole milk'.
Let's see what people who buy 'whole milk' tend to buy.  
We can see people who buy 'curd' and 'whole milk' tend to buy 'yogurt', too. They might be dairy products loves.

```{r}
submilk = subset(sub1, lhs %in% 'whole milk')
inspect(sort(submilk, by = "lift")[1:5])
```

